Chasing \rainbows \hspace{1pt} with the Habitable Worlds Observatory

### 2023-02-24 src/tex/ms.tex
[showyourwork] overleaf sync

\begin{abstract}
NASA recently announced the Habitable Worlds Observatory, which would detect rocky planets in the habitable zone, establish their habitability, and search them for biosignatures. 
%Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal. After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties. 
% JMa I've seen someone commented my text. I'm ok with that but I think it would be important to have a 1 sentence description of the 
Surface liquid water is central to the definition of planetary habitability. Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects. Direct imaging missions are optimised to detect planets near quadrature, which may obscure some of the phases at which these features are strongest. The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle. Planets that orbit close to edge-on have accessible phase angles limited by obscuration by the coronagraph. We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in XX systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in YY systems. The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with IWA.

>>>>

\begin{abstract}
NASA recently announced the Habitable Worlds Observatory, a coronagraphic mission which would detect rocky planets in the habitable zone, establish their habitability, and search them for biosignatures. 
%Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal. After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties. 
Surface liquid water is central to the definition of planetary habitability. Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects. Direct imaging missions are optimised to detect planets near quadrature, which may obscure some of the phases at which these features are strongest. The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle. Planets that orbit close to edge-on have accessible phase angles limited by obscuration by the coronagraph. We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in XX systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in YY systems. The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with IWA.
